GxHash: A High-Throughput, Non-Cryptographic Hashing Algorithm Leveraging Modern CPU Capabilities

### 2023-11-25 article/article.tex
Fix spelling and grammar mistakes (#19) (#26)
Co-authored-by: Ian Qvist <Genbox@users.noreply.github.com>

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for Random dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(4)            & 0,0241\% & 0,001094  & 0,000002 & 0,00034 \\
UInt32 GxHash(64)           & 0,0115\% & 0,000795  & 0,000002 & 0,00002 \\
UInt32 GxHash(1000)         & 0,0108\%  & 0,001007  & 0,000002 & 0,00007 \\
UInt32 HighwayHash(4)        & 0,0237\% & 0,001135  & 0,000002 & 0,00001 \\
UInt32 HighwayHash(64)       & 0,0117\% & 0,001028  & 0,000002 & 0,00078 \\
UInt32 HighwayHash(1000)     & 0,0103\% & 0,00092   & 0,000002 & 0,00032 \\
UInt32 T1ha(4)               & 0,021\%  & 0,001034  & 0,000002 & 0,00002 \\
UInt32 T1ha(64)              & 0,0123\% & 0,000933  & 0,000002 & 0,00047 \\
UInt32 T1ha(1000)            & 0,0114\% & 0,001087  & 0,000002 & 0,00027 \\
UInt32 XxHash(4)            & 0,0119\% & 0,00102   & 0,000002 & 0,00027 \\
UInt32 XxHash(64)           & 0,013\%  & 0,000871  & 0,000002 & 0,00083 \\
UInt32 XxHash(1000)         & 0,0131\% & 0,001214  & 0,000002 & 0,00038 \\
UInt32 Fnv1a(4)             & 0,031\%  & 0,001008  & 0,000002 & \textcolor{red}{0,20155} \\
UInt32 Fnv1a(64)            & 0,0094\% & 0,000748  & 0,000002 & \textcolor{orange}{0,08599} \\
UInt32 Fnv1a(1000)          & 0,0138\% & 0,000821  & 0,000002 & \textcolor{orange}{0,07861} \\
UInt32 Crc(4)               & 0,0119\% & 0,000811  & 0,000002 & \textcolor{red}{0,11689} \\
UInt32 Crc(64)              & 0,0117\% & 0,001041  & 0,000002 & \textcolor{orange}{0,02473} \\
UInt32 Crc(1000)            & 0,0123\% & 0,001097  & 0,000002 & 0,00514 \\
\hline
\end{tabular}
\caption{Quality benchmark results for the random dataset at 1,000,000 iterations}
\label{tab:quality-data-random}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for Sequential dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(4)                & 0,013\% & 0,001132  & 0,000002 & 0,00044 \\
UInt32 GxHash(64)               & 0,0112\% & 0,000651  & 0,0000019 & 0,00039 \\
UInt32 GxHash(1000)             & 0,0126\% & 0,000818  & 0,000002 & 0,00029 \\
UInt32 HighwayHash(4)            & 0,0117\% & 0,00112   & 0,000002 & 0,00011 \\
UInt32 HighwayHash(64)           & 0,0104\% & 0,001204  & 0,000002 & 0,00044 \\
UInt32 HighwayHash(1000)         & 0,0112\% & 0,001188  & 0,000002 & 0,00131 \\
UInt32 T1ha(4)                   & 0,012\%  & 0,000746  & 0,000002 & 0,00076 \\
UInt32 T1ha(64)                  & 0,0125\% & 0,000987  & 0,000002 & 0,00071 \\
UInt32 T1ha(1000)                & 0,0113\% & 0,000944  & 0,000002 & 0,00003 \\
UInt32 XxHash(4)                & 0\%      & 0,000933  & 0,000002 & 0,00018 \\
UInt32 XxHash(64)               & 0\%      & 0,000907  & 0,000002 & 0,00046 \\
UInt32 XxHash(1000)             & 0\%      & 0,001081  & 0,000002 & 0,0007 \\
UInt32 Fnv1a(4)                 & 0\%      & 0,00009   & 0,0000017 & \textcolor{red}{0,18255} \\
UInt32 Fnv1a(64)                & 0\%      & 0,000064  & 0,0000022 & \textcolor{orange}{0,08281} \\
UInt32 Fnv1a(1000)              & 0\%      & 0,000042  & 0,0000018 & \textcolor{orange}{0,08416} \\
UInt32 Crc(4)                   & 0\%      & 0,000003  & 0,000002 & \textcolor{red}{0,11729} \\
UInt32 Crc(64)                  & 0\%      & 0,000003  & 0,0000004 & \textcolor{orange}{0,02542} \\
UInt32 Crc(1000)                & 0\%      & 0,00001   & 0,0000004 & 0,0046 \\
\hline
\end{tabular}
\caption{Quality benchmark results for the sequential dataset at 1,000,000 iterations}
\label{tab:my_label}
\end{table}

@@

TODO
Regarding distribution, we can notice that GxHash outperforms HighwayHash\cite{highwayhash}, XxHash\cite{xxhash}, and T1ha0\cite{rust-t1ha}. Avalanche is slightly worse, however, possibly due to the tradeoff of doing fewer operations for greater performances. Overall, the numbers are all still very low and remain in the same ballpark, except for FNV-1a and CRC which still suffer from a relatively "high" avalanche.

@@

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for MarkovWords dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(64)                & 0,0108\% & 0,00122  & 0,000002 & 0,00044 \\
UInt32 GxHash(1000)              & 0,0121\% & 0,001106  & 0,000002 & 0,00035 \\
UInt32 HighwayHash(64)            & 0,0123\% & 0,000809  & 0,000002 & 0,00048 \\
UInt32 HighwayHash(1000)          & 0,0117\% & 0,00092   & 0,000002 & 0,00078 \\
UInt32 T1ha(64)                   & 0,0111\% & 0,000803  & 0,000002 & 0,00064 \\
UInt32 T1ha(1000)                 & 0,0123\% & 0,001175  & 0,000002 & 0,00135 \\
UInt32 XxHash(64)                & 0,0106\% & 0,000766  & 0,000002 & 0,00046 \\
UInt32 XxHash(1000)              & 0,01\%   & 0,000892  & 0,000002 & 0,00021 \\
UInt32 Fnv1a(64)                 & 0,0122\% & 0,000998  & 0,000002 & \textcolor{orange}{0,08585} \\
UInt32 Fnv1a(1000)               & 0,0127\% & 0,000993  & 0,000002 & \textcolor{orange}{0,08143} \\
UInt32 Crc(64)                   & 0,0124\% & 0,000965  & 0,000002 & \textcolor{orange}{0,02467} \\
UInt32 Crc(1000)                 & 0,0123\% & 0,000708  & 0,000002 & 0,00499 \\
\hline
\end{tabular}
\caption{Quality benchmark results for words dataset at 1,000,000 iterations}
\label{tab:my_label}
\end{table}

@@

\begin{figure}[H]
\centering
\includegraphics[width=1\textwidth]{quality-markov.png}
\caption{Distribution map for words dataset}
\label{fig:quality-sequential}
\end{figure}

>>>>

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for Random dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(4)            & 0,0241\% & 0,001094  & 0,000002 & 0,00034 \\
UInt32 GxHash(64)           & 0,0115\% & 0,000795  & 0,000002 & 0,00002 \\
UInt32 GxHash(1000)         & 0,0108\%  & 0,001007  & 0,000002 & 0,00007 \\
UInt32 HighwayHash(4)        & 0,0237\% & 0,001135  & 0,000002 & 0,00001 \\
UInt32 HighwayHash(64)       & 0,0117\% & 0,001028  & 0,000002 & 0,00078 \\
UInt32 HighwayHash(1000)     & 0,0103\% & 0,00092   & 0,000002 & 0,00032 \\
UInt32 T1ha(4)               & 0,021\%  & 0,001034  & 0,000002 & 0,00002 \\
UInt32 T1ha(64)              & 0,0123\% & 0,000933  & 0,000002 & 0,00047 \\
UInt32 T1ha(1000)            & 0,0114\% & 0,001087  & 0,000002 & 0,00027 \\
UInt32 XxHash(4)            & 0,0119\% & 0,00102   & 0,000002 & 0,00027 \\
UInt32 XxHash(64)           & 0,013\%  & 0,000871  & 0,000002 & 0,00083 \\
UInt32 XxHash(1000)         & 0,0131\% & 0,001214  & 0,000002 & 0,00038 \\
UInt32 Fnv1a(4)             & 0,031\%  & 0,001008  & 0,000002 & \textcolor{red}{0,20155} \\
UInt32 Fnv1a(64)            & 0,0094\% & 0,000748  & 0,000002 & \textcolor{orange}{0,08599} \\
UInt32 Fnv1a(1000)          & 0,0138\% & 0,000821  & 0,000002 & \textcolor{orange}{0,07861} \\
UInt32 Crc(4)               & 0,0119\% & 0,000811  & 0,000002 & \textcolor{red}{0,11689} \\
UInt32 Crc(64)              & 0,0117\% & 0,001041  & 0,000002 & \textcolor{orange}{0,02473} \\
UInt32 Crc(1000)            & 0,0123\% & 0,001097  & 0,000002 & 0,00514 \\
\hline
\end{tabular}
\caption{Quality benchmark results for the random dataset at 1,000,000 iterations}
\label{tab:scores-random}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for Sequential dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(4)                & 0,013\% & 0,001132  & 0,000002 & 0,00044 \\
UInt32 GxHash(64)               & 0,0112\% & 0,000651  & 0,0000019 & 0,00039 \\
UInt32 GxHash(1000)             & 0,0126\% & 0,000818  & 0,000002 & 0,00029 \\
UInt32 HighwayHash(4)            & 0,0117\% & 0,00112   & 0,000002 & 0,00011 \\
UInt32 HighwayHash(64)           & 0,0104\% & 0,001204  & 0,000002 & 0,00044 \\
UInt32 HighwayHash(1000)         & 0,0112\% & 0,001188  & 0,000002 & 0,00131 \\
UInt32 T1ha(4)                   & 0,012\%  & 0,000746  & 0,000002 & 0,00076 \\
UInt32 T1ha(64)                  & 0,0125\% & 0,000987  & 0,000002 & 0,00071 \\
UInt32 T1ha(1000)                & 0,0113\% & 0,000944  & 0,000002 & 0,00003 \\
UInt32 XxHash(4)                & 0\%      & 0,000933  & 0,000002 & 0,00018 \\
UInt32 XxHash(64)               & 0\%      & 0,000907  & 0,000002 & 0,00046 \\
UInt32 XxHash(1000)             & 0\%      & 0,001081  & 0,000002 & 0,0007 \\
UInt32 Fnv1a(4)                 & 0\%      & 0,00009   & 0,0000017 & \textcolor{red}{0,18255} \\
UInt32 Fnv1a(64)                & 0\%      & 0,000064  & 0,0000022 & \textcolor{orange}{0,08281} \\
UInt32 Fnv1a(1000)              & 0\%      & 0,000042  & 0,0000018 & \textcolor{orange}{0,08416} \\
UInt32 Crc(4)                   & 0\%      & 0,000003  & 0,000002 & \textcolor{red}{0,11729} \\
UInt32 Crc(64)                  & 0\%      & 0,000003  & 0,0000004 & \textcolor{orange}{0,02542} \\
UInt32 Crc(1000)                & 0\%      & 0,00001   & 0,0000004 & 0,0046 \\
\hline
\end{tabular}
\caption{Quality benchmark results for the sequential dataset at 1,000,000 iterations}
\label{tab:scores-sequential}
\end{table}

@@

Overall, the distribution/avalanche scores are all very low and remain in the same ballpark for GxHash, HighwayHash, T1ha and XxHash. FNV-1a and CRC have relatively worse avalanche score due to their simplicity.

@@

\begin{table}[H]
\centering
\begin{tabular}{|l|r|r|r|r|}
\hline
\textbf{Function for MarkovWords dataset} & \textbf{Collisions} & \textbf{Bits Distribution} & \textbf{Distribution} & \textbf{Avalanche} \\
\hline
UInt32 GxHash(64)                & 0,0108\% & 0,00122  & 0,000002 & 0,00044 \\
UInt32 GxHash(1000)              & 0,0121\% & 0,001106  & 0,000002 & 0,00035 \\
UInt32 HighwayHash(64)            & 0,0123\% & 0,000809  & 0,000002 & 0,00048 \\
UInt32 HighwayHash(1000)          & 0,0117\% & 0,00092   & 0,000002 & 0,00078 \\
UInt32 T1ha(64)                   & 0,0111\% & 0,000803  & 0,000002 & 0,00064 \\
UInt32 T1ha(1000)                 & 0,0123\% & 0,001175  & 0,000002 & 0,00135 \\
UInt32 XxHash(64)                & 0,0106\% & 0,000766  & 0,000002 & 0,00046 \\
UInt32 XxHash(1000)              & 0,01\%   & 0,000892  & 0,000002 & 0,00021 \\
UInt32 Fnv1a(64)                 & 0,0122\% & 0,000998  & 0,000002 & \textcolor{orange}{0,08585} \\
UInt32 Fnv1a(1000)               & 0,0127\% & 0,000993  & 0,000002 & \textcolor{orange}{0,08143} \\
UInt32 Crc(64)                   & 0,0124\% & 0,000965  & 0,000002 & \textcolor{orange}{0,02467} \\
UInt32 Crc(1000)                 & 0,0123\% & 0,000708  & 0,000002 & 0,00499 \\
\hline
\end{tabular}
\caption{Quality benchmark results for words dataset at 1,000,000 iterations}
\label{tab:scores-words}
\end{table}

@@

\begin{figure}[H]
\centering
\includegraphics[width=1\textwidth]{quality-markov.png}
\caption{Distribution map for words dataset}
\label{fig:quality-words}
\end{figure}
